---
title: "Reading Delimited Data"
format: html
toc: TRUE
editor_options: 
  chunk_output_type: inline
---

## Glass Data

1.  
installing (commented out so it doesn't rerun each time) and loading the tidyverse package to read in the dataset, then naming each column within the read_csv function.
```{r}
#install.packages("tidyverse")
library(tidyverse)
Glass <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type_of_glass"))
```

2. 
using the mutate and factor function to create a chain that renames the numeric values of Type_of_glass to descriptive character strings.
```{r}
Glass |> 
  mutate(Type_of_glass = Type_of_glass |> 
           factor(levels = 1:7, labels = c("building_windows_float_processed", "building_windows_non_float_processed", "vehicle_windows_float_processed", "vehicle_windows_non_float_processed", "containers", "tableware", "headlamps")))
```

3. 
Continuing the chain and using the filter function to keep Fe variables whose value is less than 0.2 as well as tableware and headlamp values for Type_of_glass.
```{r}
Glass <- Glass |> 
  mutate(Type_of_glass = Type_of_glass |>
           factor(levels = 1:7, labels = c("building_windows_float_processed", "building_windows_non_float_processed", "vehicle_windows_float_processed", "vehicle_windows_non_float_processed", "containers", "tableware", "headlamps"))) |> 
  filter(Fe < 0.2) |>
  filter(Type_of_glass == c("headlamps", "tableware"))
Glass
```

## Yeast Data

1. 
Reading Yeast data into R and naming the columns
```{r}
Yeast <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class"))
Yeast
```

2. 
beginning of the chain which removes columns seq_name and nuc.
```{r}
Yeast |>
  select(-seq_name, -nuc)
```

3. 
Continuing the chain to include columns corresponding to the mean and median of each variable at each class grouping using the mutate and across function.
```{r}
Yeast |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(across(where(is.numeric), list(mean = mean, median = median), names = "{.col}_{.fn}"))
```
